Tool-driven Revolutions in Archaeological Science

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
added Stupler_2017 to references
reordered part of discussion so we don't jump from data to code back to data.

Rather, this change enables archaologists to more faithfully adhere to the fundamental tenet of the scientific process which demands the reproducibility of results. We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages. But instead that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice. By using and sharing code we can undo this drift, and return to a robust scrietific practice for archaeology.

There is a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses. In the digital age, though, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses. This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results. Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses. As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences. These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013].

>>>>

Rather, this change enables archaologists to more faithfully adhere to the fundamental tenet of the scientific process which demands the reproducibility of results. We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages. There is, for example, a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses. Archaeological research has always relied on the reuse of data [@Strupler_2017].

Nonetheless we feel that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice. In the digital age, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses. This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results. Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses. By using and sharing code we can undo this drift, and return to a robust scrietific practice for archaeology. As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences. These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013]. Also in archaeology, more and more funding agencies push for open science methods and transparency [@Stupler_2017].
